---
title: "first_glance"
bibliography: boot-camp.bib
csl: poetics.csl
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Scientific Questions

Routing algorithms in the ocean are hard [@christiansen_ship_2004].

What kind of questions and results are we looking for?

- Q: How can we improve routing from a simple grid?
  
  - A: [Generic] Use triangulated network. Compare distances between start/end points. Also compare computational time.
  
  - A: [Application] Compare with extra whale risk conserved & less shipping distance traversed for this example. Results:
  
    - Show boxplot of grid vs triangulated routing and efficiencies gained on y axis for whales & shipping
    
    - Show plot of how parameter for triangulation (like angle) on x-axis changes, how efficincies are gained in whale conservation / shipping distance on y-axis 
    
    - Showing the Shiny app before/after this new routing mechanism. Embedded as a HTML slideshow.

- Q: How can we get faster results for the shortest path problem than Dijkstra's algorithm?

 - A: [Generic] Use the A* algorithm, or one of its variants depending on how large the graph is. We can create large sample graphs to compare times and solutions for different-sized graphs.
 
 - A: [Generic] Look into UAV path-planning methodologies, since many of those applications have continuous paths rather than discrete.


## Technical Questions

Let's try to use <https://github.com/dgrimsman/boot-camp/issues>.

## Data

British Columbia (BC)

- land: `bc_spp_gcs.shp`
- whales: `v72zw_epsg3857.grd`

### raster of whale hot spots

```{r whales, message=F}
library(raster) # install.packages('rgdal')

whales = raster('data/v72zw_epsg3857.grd')
plot(whales)
```

### ports to route between

```{r ports, message=FALSE}
library(readr)

ports = read_csv('data/ports_bc.csv')
```


### interactive map

```{r}
library(leaflet)

leaflet() %>%
  addProviderTiles("Stamen.TonerLite", options = providerTileOptions(noWrap = TRUE)) %>% 
  addRasterImage(
    whales, opacity = 0.8, project = F, group='Raster',
    colors = colorNumeric(
      palette = 'Reds', domain = c(cellStats(whales, 'min'), cellStats(whales, 'max')), 
      na.color = "#00000000", alpha = TRUE)) %>%
  addCircleMarkers(
    ~lon, ~lat, color='blue', data=ports, layerId=~name, group='Ports',
    popup = ~sprintf('<b>%s</b><br>%0.2f, %0.2f', name, lon, lat)) %>%
  addLayersControl(
    overlayGroups = c('Raster', 'Ports'),
    options = layersControlOptions(collapsed=T))
```

## Testing New Packages

### RBGL

```{r install, eval=FALSE}
source("https://bioconductor.org/biocLite.R")
biocLite("RBGL")
```

```{r RBGL not working, eval=F}
library(RBGL)

# astar search
con <- file(system.file("XML/dijkex.gxl",package="RBGL"), open="r")
coex <- fromGXL(con)
close(con)
astarSearch(coex)
```

### RTriangle

- [Computational Geometry in Python](http://blancosilva.github.io/post/2014/10/28/Computational-Geometry-in-Python.html#convex-hulls)

- [Triangle: .poly files](http://www.cs.cmu.edu/~quake/triangle.poly.html)

```{r}
library(RTriangle)
library(rgdal)
library(rgeos)
library(RTriangle)
library(dplyr)
library(Matrix)
library(igraph)
library(sp)
library(gdistance)
triangulate = RTriangle::triangulate # override rgeos::triangulate

## Create an object with a concavity
p <- pslg(P=rbind(c(0, 0), c(0, 1), c(0.5, 0.5), c(1, 1), c(1, 0)),
          S=rbind(c(1, 2), c(2, 3), c(3, 4), c(4, 5), c(5, 1)))

## Plot it
plot(p)

## Triangulate it
tp <- triangulate(p)
plot(tp)

## Load a data set containing a hole
A <- read.pslg(file.path(system.file(package = "RTriangle"), "extdata", "A.poly"))
plot(A)
## Triangulate the PSLG
tA <- triangulate(A)
plot(tA)
## Triangulate the PSLG with triangles in which no angle
## is smaller than 20 degrees
tA <- triangulate(A, q=20)
plot(tA)
## Triangulate the PSLG with triangles in which no triangle has 
## area greater than 0.001
tA <- triangulate(A, a=0.001)
plot(tA)

source('custom.R')
ply = read.pslg('data/superior.poly')
plot(ply)
t_ply <- triangulate(ply)
plot(t_ply)
t_ply <- triangulate(ply, q=20, D=T) # 'pq20D'
plot(t_ply)
```

## Triangulate BC

[Triangularize BC raster · Issue #1 · dgrimsman/boot-camp](https://github.com/dgrimsman/boot-camp/issues/1)

### Get Extent Polygon of Raster

```{r create bc.geojson, eval=FALSE}
whales_1 = whales
whales_1[!is.na(whales_1)] = 1
plot(whales_1)
whales_p = rasterToPolygons(whales_1, n=8, na.rm=T, digits=12, dissolve=T)
plot(whales_p)

writeOGR(whales)
writeOGR(whales_p, 'data/bc', layer='layer', driver='GeoJSON')
file.rename('data/bc','data/bc.geojson')
```

```{r bc.poly pt 1, eval=F}
whales_p = readOGR(dsn='data/bc.geojson', layer='OGRGeoJSON')

#names(whales_p)
#slotNames(whales_p)
#whales_p@polygons

# vertices = data_frame(i=numeric(), x=numeric(), y=numeric())
# for (i in 1:length(whales_p@polygons[[1]]@Polygons)){ # p = whales_p@polygons[[1]]@Polygons[1]
#   vertices = bind_rows(
#     vertices,
#     a = whales_p@polygons[[1]]@Polygons[[i]]@coords %>%
#       as.data.frame() %>%
#       mutate(i = i))
# }

vertices = data_frame(i=numeric(), x=numeric(), y=numeric())
segments = data_frame(i=numeric(), start=numeric(), fin=numeric())
start_v = 1
fin_v = 1
for (i in 1:length(whales_p@polygons[[1]]@Polygons)){ # p = whales_p@polygons[[1]]@Polygons[1]
  p = whales_p@polygons[[1]]@Polygons[[i]]
  if (p@area < 1e10){
    next
  }
  
  # save vertices
  coords = rename(as.data.frame(p@coords), c('V1'='x', 'V2'='y'))
  vertices = bind_rows(vertices, coords)
  vertices = head(vertices, -1) # because last vertex is redundant
  
  #save segments
  num_vert = dim(p@coords)[1] - 1
  fin_v = start_v + num_vert - 1
  start = c(start_v:fin_v)
  fin = c((start_v + 1):fin_v, start_v)
  segments = bind_rows(
    segments,
    a = data_frame(start, fin))
  start_v = fin_v + 1
}
vertices$i<-seq.int(nrow(vertices))
segments$i<-seq.int(nrow(segments))

idx = which(duplicated(vertices[,c('x','y')]))
vertices$x[idx] = vertices$y[idx] + runif(length(idx), min=0.01, max=0.03)
```

```{r bc.poly pt 2, eval=F}
# http://www.cs.cmu.edu/~quake/triangle.poly.html
# DONE: @bbest
# First line: <# of vertices> <dimension (must be 2)> <# of attributes> <# of boundary markers (0 or 1)>
# Following lines: <vertex #> <x> <y> [attributes] [boundary marker]
poly = 'data/bc.poly'
write_file(sprintf('%d 2 0 0\n', nrow(vertices)), poly)
for (i in 1:nrow(vertices)){ # i =1
  row = vertices[i,]
  write_file(sprintf('%d %0.2f %0.2f\n', i, row$x, row$y), poly, append=T)
}

# DONE: @dgrimsman
# One line: <# of segments> <# of boundary markers (0 or 1)>
# Following lines: <segment #> <endpoint> <endpoint> [boundary marker]
write_file(sprintf('%d 0\n', nrow(segments)), poly, append=T)
for (i in 1:nrow(segments)){ # i =1
  row = segments[i,]
  write_file(sprintf('%d %g %g\n', i, row$start, row$fin), poly, append=T)
}

# TODO: @dgrimsman
# One line: <# of holes>
# Following lines: <hole #> <x> <y>
# # holes = data_frame(i=numeric(), x=numeric(), y=numeric())
# write_file(sprintf('%d\n', length(whales_p@polygons[[1]]@Polygons) - 1), poly, append=T)
# j = 1
# for (i in 1:length(whales_p@polygons[[1]]@Polygons)){
#   p = whales_p@polygons[[1]]@Polygons[[i]]
#   if (p@area <1e10){
#     hole_coord = p@labpt
#     x = hole_coord[1]
#     y = hole_coord[2]
#     # holes = rbind(holes, c(i, hole_coord[1, 1], hole_coord[1, 2]))
#     write_file(sprintf('%d %g %g\n', j, x, y), poly, append=T)
#     j = j + 1
#   }
# }

write_file(sprintf('%d\n', 0), poly, append=T)
```

```{r triangulate}
library(RTriangle)
source('custom.R')
bc_poly = read.pslg('data/bc.poly') # not yet working b/c segments needed
plot(bc_poly)
bc_poly_tri <- triangulate(bc_poly, q=20, D=T) # 'pq20D'
plot(bc_poly_tri)
```

```{r shortest}
library(igraph)
library(Matrix)

adj_mat_txt = 'data/adj_mat.txt'

if (!file.exists(adj_mat_txt)){
  
  # create adjacency matrix
  num_v_tri = dim(bc_poly_tri$P)[1]
  adj_mat <- Matrix(0, nrow = num_v_tri, ncol = num_v_tri, sparse=TRUE)
  for (i in 1:dim(bc_poly_tri$E)[1]){
    edge = bc_poly_tri$E[i,]
    p1 = edge[1]
    p2 = edge[2]
    d = sqrt((bc_poly_tri$P[p1, 1] - bc_poly_tri$P[p2, 1])^2 + (bc_poly_tri$P[p1, 2] - bc_poly_tri$P[p2, 2])^2)
    adj_mat[p1, p2] = d
    adj_mat[p2, p1] = d
  }
  writeMM(adj_mat,file=adj_mat_txt)
  
} else {
  
  # read existing
  readMM(adj_mat_txt)
}

library(sp)
library(raster)
lonlat1 = SpatialPoints(data.frame(-127.08, 50.59))
lonlat2 = SpatialPoints(data.frame(-130.32, 54.31))

shortest_path_tri = function(lonlat1, lonlat2){

  xy = c(coordinates(lonlat1), coordinates(lonlat2)) %>%
        matrix(ncol=2, byrow=T) %>%
        SpatialPoints(crs("+init=epsg:4326")) %>%
        spTransform(crs("+init=epsg:3857"))
  
  #create graph and find shortest path
  graph = graph_from_adjacency_matrix(adj_mat, weighted=TRUE, mode="undirected")
  pts_tri = plyr::rename(as.data.frame(bc_poly_tri$P), c('V1'='x', 'V2'='y'))
  pts_tri$dist_start = sqrt((pts_tri$x - xy@coords[1,1])^2 + (pts_tri$y - xy@coords[1,2])^2)
  pts_tri$dist_end = sqrt((pts_tri$x - xy@coords[2,1])^2 + (pts_tri$y - xy@coords[2,2])^2)
  start_pt = as.numeric(rownames(pts_tri[order(pts_tri$dist_start),])[1])
  end_pt = as.numeric(rownames(pts_tri[order(pts_tri$dist_end),])[1])
  path_tri = shortest_paths(graph, start_pt, end_pt)
  return(path_tri)
}

path_tri2sl = function(path){
  # cast the path as a SpatialLines object
  ln_xy = data_frame(x=numeric(), y=numeric())
  for (i in 1:length(path$vpath[[1]])){
    pt = as.numeric(path$vpath[[1]][i])
    # bind node coordinates into row of line
    ln_xy = rbind(
      ln_xy,
      data_frame(
        x=bc_poly_tri$P[pt,1], 
        y=bc_poly_tri$P[pt,2]))
  }
  return(SpatialLines(list(Lines(Line(ln_xy), ID="a"))))
}

# iterate through nodes in path
get_dist_tri = function(path){
  dist_tri = 0
  prev_pt = as.numeric(path$vpath[[1]][1])
  for (i in 2:length(path$vpath[[1]])){ # i=2
    pt = as.numeric(path$vpath[[1]][i])
    
    # update distance from adjacency matrix
    dist_tri = dist_tri + adj_mat[pt, prev_pt]
    
    # cat(sprintf('%d -> %d: %g\n', pt, prev_pt, adj_mat[pt, prev_pt]))
    prev_pt = pt
  }
  #c at(sprintf('TOTAL length: %g\n', dist_tri))
  return(dist_tri)
}
```

```{r eval=F}
# ones map for getting linear path
r1 = whales
r1[!is.na(whales)] = 1

shortest_path_grid = function(lonlat1, lonlat2){
  xy = c(coordinates(lonlat1), coordinates(lonlat2)) %>%
          matrix(ncol=2, byrow=T) %>%
          SpatialPoints(crs("+init=epsg:4326")) %>%
          spTransform(crs("+init=epsg:3857"))
  
  # update to nearest non-NA points on raster for shortestPath to work
  xy = c(
    coordinates(r1)[which.min(mask(distanceFromPoints(r1, xy[1]), r1)),],
    coordinates(r1)[which.min(mask(distanceFromPoints(r1, xy[2]), r1)),]) %>%
    matrix(ncol=2, byrow=T) %>%
    SpatialPoints(crs("+init=epsg:3857"))
  
  path_grid = shortestPath(
      geoCorrection(transition(1 / (r1), mean, directions=8), type="c"), 
      xy[1],
      xy[2],
      output='SpatialLines')
  return(path_grid)
}

#get the length of the path
get_dist_grid = function(path){
  coords = path@lines[[1]]@Lines[[1]]@coords
  dist_grid = 0
  prev_pt = coords[1,]
  for (i in 2:dim(coords)[1]){
    pt = coords[i,]
    dist_grid = dist_grid + sqrt((pt[1] - prev_pt[1])^2 + (pt[2] - prev_pt[2])^2)
    prev_pt = pt
  }
  return(dist_grid)
}
```


## Results


- plot for single path comparing raster vs vector paths, 2 panels (L, R):
  - Left: raster network with path (nodes = #, dist = #)
  - Right: triangulated network with path (nodes = #, dist = #)
  - We're using fewer nodes to extract shorter paths.

```{r}
# zoom
plot(raster::rasterToPoints(r1), pch=20, xlim=-14)

plot(whales)

plot(whales, xlim = c(-14257523, -14121332), ylim = c(6547005, 6647474))

labels = c('South Entrance', 'McNeill', 'Prince Rupert', 'Kitimat', 'Texada Island', 'Crofton')
lons = c(-131, -127.08, -130.32, -128.67, -124.41, -123.65)
lats = c(52, 50.59, 54.31, 54.01, 49.66, 48.86)
my_ports = data.frame(labels, lons, lats)
data = data.frame(start=character(), end=character(), dist_grid=numeric(), dist_tri=numeric())

for (i in 1:dim(my_ports)[1]){
  for (j in i+1:dim(my_ports)[1]){ 
    lonlat1 = SpatialPoints(my_ports[i, 2:3])
    lonlat2 = SpatialPoints(my_ports[j, 2:3])
    
    path_grid = shortest_path_grid(lonlat1, lonlat2)
    dist_grid = get_dist_grid(path_grid)
    
    #dir.create('plots')
    #png('plots/grid-path_zoom.png')
    #plot(raster::rasterToPoints(r1), pch=20) #, xlim = c(-14257523, -14121332), ylim = c(6547005, 6647474), cex=0.2)
    #plot(path_grid, col='green', lwd=5, add=T)
    #legend('topright', sprintf('n=%s\nd=%s', scales::comma(cellStats(r1, 'sum')), scales::comma(dist_grid)))
    #dev.off()
    
    spt = shortest_path_tri(lonlat1, lonlat2)
    dist_tri = get_dist_tri(spt)
    path_tri_ln = path_tri2sl(spt)
    
    png(sprintf('plots/paths_from_%s_to_%s.png', my_ports[i, 1], my_ports[j, 1]))
    plot(bc_poly_tri)# , xlim = c(-14257523, -14121332), ylim = c(6547005, 6647474))
    plot(path_tri_ln, col='blue', lwd=5, add=T)
    plot(path_grid, col='green', lwd=5, add=T)
    dev.off()
    #legend('topright', sprintf('# Nodes = %s\nPath Length = %d km\n', scales::comma(dim(bc_poly_tri$P)[1]), round(dist_tri/1000)))
  }
}
write.table(data, file='data/path_lengths')
```


  
- run between different ports: get difference b/n methods & order -/+, then isolate a few of the outlier paths in either direction

- table of computational time

## R Tips Background

- [HTML Documents - rmarkdown](http://rmarkdown.rstudio.com/html_document_format.html#code_folding)
- [Bibliographies and Citations](http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html)
- [bbest/rmarkdown-example: R Markdown example showing figures & tables with captions, equations, inline R values and references with a Zotero library](https://github.com/bbest/rmarkdown-example)


## References

